Об оценках на диаметр плоских множеств с целочисленными расстояниями полуобщего положения \footnote{Исследование выполнено за счет гранта Российского научного фонда (проект 19-11-00197).}

### 2019-11-01 semi-general-suplinear-bound/CHEBSB_Avdeev.tex
ChebSb bibliography - all but DOIs
Self-proofreading iteration

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% {{ biblatex
%% Команда \Cite (с большой буквы) у biblatex своя.
%% Сохраняем "нашу"
\let\CiteChebSb=\Cite
%% Убираем \Cite
\let\Cite=\relax
%% Вот теперь смело загружаем biblatex
\usepackage[backend=biber,style=gost-numeric,sorting=none]{biblatex}
\addbibresource{../common/notmy.bib}
\addbibresource{../common/my.bib}
%% Восстанавливаем "нашу" \Cite
\let\Cite=\CiteChebSb
%% }} biblatex

@@

\Abstract
{%Аннотация статьи на русском языке 150-250 слов с учетом ключевых слов
Множество точек $M$ на плоскости называется плоским множеством с целочисленными расстояниями,
если все расстояния между точками $M$ суть целые числа,
и при этом $M$ не содержится ни в какой прямой.
Говорят, что плоское множество с целочисленными расстояниями есть множество полуобщего положения,
если никакие три его точки не лежат на одной прямой.
Известная оценка снизу на диаметр плоского множества с целочисленными расстояниями
линейна относительно его мощности.
Ранее не были известны отдельные оценки снизу на диаметр плоских множеств с целочисленными расстояниями полуобщего положения
заданной мощности
(известная конструктивная оценка сверху на диаметр плоских множеств с целочисленными расстояниями
использует как раз множества полуобщего положения).
В статье доказывается надлинейная оценка снизу
на диаметр плоского множества с целочисленными расстояниями полуобщего положения
(полиномиальная с показателем $5/4$).
Доказательство основано на относительно большом количестве лемм и наблюдений,
включая результаты Солимоси
из статьи, в которой была впервые доказана линейная оценка снизу
на диаметр плоских множеств с целочисленными расстояниями.
Полученная вданной статье оценка всё ещё не точна.
}
{%Аннотация статьи на английском языке
A point set $M$ in the Euclidean plane is called a planar integral point set if all the distances between the
elements of $M$ are integers, and $M$ is not situated on a straight line.
A planar integral point set is called to be in semi-general position, if it does not contain collinear triples.
The existing lower bound for mininum diameter of planar integral point sets is linear on its cardinality.
There were no known special diameter bounds for planar integral point sets in semi-general position
(the known upper bound for planar integral point sets is constructive
and provides planar integral point sets in semi-general position).
We prove a new lower bound for mininum diameter of planar integral point sets in semi-general position
that is better than linear (polynomial with power $5/4$).
The proof is based on several lemmas observations, including the ones provided by Solymosi
while proving the first linear lower bound for diameter of a planar integral point set.
We have to admit that our bound is still not tight.
}

@@

There are some bounds for minimal diameter of planar integral point sets in some special positions.
Assuming that the planar integral point sets contains many collinear points,
the following result holds.
\begin{Theorem}~\cite[Theorem 4]{kurz2008minimum}
	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
	at least $n^\delta$ collinear points, there exists a $n_0 (\varepsilon)$
	such that for all $n \geq n_0 (\varepsilon)$ we have
	\begin{equation}
		\operatorname{diam} P \geq n^{\frac{\delta}{4 \log 2(1+\varepsilon)}\log \log n}
		.
	\end{equation}
\end{Theorem}
For diameter bounds for circular sets, we refer the reader to~\cite{bat2018number}.

Particular cases of planar integral point sets are also discussed
in~\cite[\S 5.11]{brass2006research},~\cite[\S D20]{guy2013unsolved},~\cite{our-pmm-2018},~\cite{our-ped-2018}.
For generalizaton in higher dimensions and the appropriate bounds, see~\cite{kurz2005characteristic,nozaki2013lower}.

@@

\begin{lemma}
	\cite[Observation 1]{solymosi2003note}
	If a triangle $T$ has integer side-lengths $a \leq b \leq c$,
	then the minimal height $m$ of it is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
\end{lemma}

@@

\begin{lemma}
	\cite{smurov1998stripcoverings}
	If a triange $T$ with minimal height $\rho$ is situated in a strip,
	then the width of a strip is at least $\rho$.
\end{lemma}

\begin{Corollary}
	\label{cor:solymosi_strip}
	If a triangle $T$ with integer side-lengths $a \leq b \leq c$ is situated in a strip,
	then the width of a strip is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
\end{Corollary}

@@

	For $m \leq p^{2/5}$, Lemma~\ref{lem:count_of_points_on_hyperbolas} yields that
	\begin{equation}
		n \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| \leq  4 p^{4/5}
		,
	\end{equation}
	or, that is the same,
	\begin{equation}
		\label{eq:hyperbolas_5_4}
		p \geq (n/4) ^ {5/4} > (n/5) ^ {5/4}
		.
	\end{equation}

@@

\section{Conclusion}
The presented bound is the first special lower bound for sets in semi-general position.
Thus, we did not accepted the challenge to make the constant in Theorem~\ref{thm:main_result} as large as possible,
in order to keep the ideas of the proof clear and understandable.
A more thorough research can be done in the future to enlarge the constant.
However, the upper and lower bounds are still not tight.

@@

\printbibliography

%библиография по ГОСТу
\begin{thebibliography}{99}

	\bibitem{1}
	Banks~W.\,D., Conflitti~A., Shparlinski~I.\,E. Character sums over integers with restricted $g$-ary digits // Illinois J. Math. 2002. Vol. 46, №3. P. 819-836.

\end{thebibliography}

%библиография по Гарвардскому стандарту
\begin{engbibliography}{99}

	\bibitem{en1}
	Banks, W.\,D., Conflitti, A. \& Shparlinski, I.\,E. 2002, ``Character sums over integers with restricted $g$-ary digits``, \textit{Illinois J. Math.}, vol. 46, no. 3, pp. 819-836.

\end{engbibliography}

>>>>

@@

\Abstract
{%Аннотация статьи на русском языке 150-250 слов с учетом ключевых слов
Множество точек $M$ на плоскости называется плоским множеством с целочисленными расстояниями,
если все расстояния между точками $M$ суть целые числа,
и при этом $M$ не содержится ни в какой прямой.
Говорят, что плоское множество с целочисленными расстояниями есть множество полуобщего положения,
если никакие три его точки не лежат на одной прямой.
Известная оценка снизу на диаметр плоского множества с целочисленными расстояниями
линейна относительно его мощности.
Ранее не были известны отдельные оценки снизу на диаметр плоских множеств с целочисленными расстояниями полуобщего положения
заданной мощности
(известная конструктивная оценка сверху на диаметр плоских множеств с целочисленными расстояниями
использует как раз множества полуобщего положения).
В статье доказывается надлинейная оценка снизу
на диаметр плоского множества с целочисленными расстояниями полуобщего положения
(полиномиальная с показателем $5/4$).
Доказательство основано на относительно большом количестве лемм и наблюдений,
включая результаты Солимоси
из статьи, в которой была впервые доказана линейная оценка снизу
на диаметр плоских множеств с целочисленными расстояниями.
Полученная в данной статье оценка всё ещё не точна.
}
{%Аннотация статьи на английском языке
A point set $M$ in the Euclidean plane is called a planar integral point set if all the distances between the
elements of $M$ are integers, and $M$ is not situated on a straight line.
A planar integral point set is called to be in semi-general position, if it does not contain collinear triples.
The existing lower bound for mininum diameter of a planar integral point set is linear with respect to its cardinality.
There were no known special diameter bounds for planar integral point sets in semi-general position
(the known upper bound for planar integral point sets is constructive
and provides planar integral point sets in semi-general position).
We prove a new lower bound for mininum diameter of planar integral point sets in semi-general position
that is better than linear (polynomial of power $5/4$).
The proof is based on several lemmas and observations, including the ones provided by Solymosi
while proving the first linear lower bound for diameter of a planar integral point set.
We have to admit that our bound is still not tight.
}

@@

There are some bounds for minimal diameter of planar integral point sets in some special positions.
Assuming that a planar integral point set contains many collinear points,
the following result holds.
\begin{Theorem}~\cite[Theorem 4]{kurz2008minimum}
	For $\delta > 0$, $\varepsilon > 0$, and $P\in\mathfrak{M}(2,n)$ with
	at least $n^\delta$ collinear points, there exists a $n_0 (\varepsilon)$
	such that for all $n \geq n_0 (\varepsilon)$ we have
	\begin{equation}
		\operatorname{diam} P \geq n^{\frac{\delta}{4 \log 2(1+\varepsilon)}\log \log n}
		.
	\end{equation}
\end{Theorem}
For diameter bounds for circular sets, we refer the reader to~\cite{bat2018number}.

Particular cases of planar integral point sets are also discussed
in~\cite[\S 5.11]{brass2006research},~\cite[\S D20]{guy2013unsolved},~\cite{our-pmm-2018},~\cite{our-ped-2018}.
For generalization to higher dimensions and the appropriate bounds, see~\cite{kurz2005characteristic,nozaki2013lower}.

@@

\begin{lemma}
	\cite[Observation 1]{solymosi2003note}
	If a triangle $T$ has integer side lengths $a \leq b \leq c$,
	then the minimal height $m$ of it is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
\end{lemma}

@@

\begin{lemma}
	\cite{smurov1998stripcoverings}
	If a triange $T$ with minimal height $\rho$ is situated in a strip,
	then the width of the strip is at least $\rho$.
\end{lemma}

\begin{Corollary}
	\label{cor:solymosi_strip}
	If a triangle $T$ with integer side lengths $a \leq b \leq c$ is situated in a strip,
	then the width of the strip is at least $\left(a - \frac{1}{4}\right)^{1/2}$.
\end{Corollary}

@@

	For $m \leq p^{2/5}$, Lemma~\ref{lem:count_of_points_on_hyperbolas} yields that
	\begin{equation}
		n \leq 4 \cdot |M_1 M_2| \cdot |M_3 M_4| \leq  4 p^{4/5}
		.
	\end{equation}
	This means that
	\begin{equation}
		\label{eq:hyperbolas_5_4}
		p \geq (n/4) ^ {5/4} > (n/5) ^ {5/4}
		.
	\end{equation}

@@

\section{Conclusion}
The presented bound is the first special lower bound for sets in semi-general position.
Thus, we have not accepted the challenge to make the constant in Theorem~\ref{thm:main_result} as large as possible,
in order to keep the ideas of the proof clear and understandable.
A more thorough research can be done in the future to enlarge the constant.
However, the upper and lower bounds are still not tight.

@@

%библиография по ГОСТу

% Мне библиографию в требуемом формате сгенерировал biblatex:
% https://github.com/odomanov/biblatex-gost/issues/20

\input{blg-GOST}

%библиография по Гарвардскому стандарту

\input{blg-Harvard}
